\documentclass[12pt]{iopart}

\usepackage[activate={true, nocompatibility}, final, tracking=true, kerning=true, spacing=true, factor=1100, stretch=10, shrink=10]{microtype}
\microtypecontext{spacing=nonfrench}

\usepackage[pdftex]{graphicx}
\usepackage[caption=false]{subfig}

\usepackage{siunitx}
\usepackage[ruled, linesnumbered]{algorithm2e}
\DontPrintSemicolon

\input{glossary}

\usepackage[style=authoryear, backend=biber, mincitenames=1, maxcitenames=2, minbibnames=1, maxbibnames=99, uniquelist=false, uniquename=false]{biblatex}
\addbibresource{./Biblio.bib}

\begin{document}
    \title[Data driven surrogate signal extraction for dynamic PET using selective PCA]{Data driven surrogate signal extraction for dynamic PET using selective PCA: time windows versus the combination of components}
    
    \author{Alexander~C~Whitehead$^{1, 2, 3}$, Kuan-Hao~Su$^4$, Elise~C~Emond$^1$, Ander~Biguri$^5$, Ludovica~Brusaferri$^6$, Maria~Machado$^1$, Joanna~C~Porter$^7$, Helen~Garthwaite$^7$, Scott~D~Wollenweber$^4$, Jamie~R~McClelland$^2$ and Kris~Thielemans$^{1, 2}$}
    
    \address{$^1$ Institute of Nuclear Medicine, University College London, UK}
    \address{$^2$ Centre for Medical Image Computing, University College London, UK}
    \address{$^3$ Department of Computer Science, University College London, UK}
    \address{$^4$ Molecular Imaging and Computed Tomography Engineering, GE Healthcare, USA}
    \address{$^5$ Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK}
    \address{$^6$ Computer Science and Informatics, London South Bank University, UK}
    \address{$^7$ Centre for Respiratory Medicine, University College London, UK}
    
    \ead{alexander.whitehead.18@ucl.ac.uk}
    
    \begin{abstract}
        \paragraph{Objective} Respiratory motion correction is beneficial in PET, as it can reduce artefacts caused by motion and improve quantitative accuracy. Methods of motion correction are commonly based on a respiratory trace obtained through an external device (like the Real Time Position Management System) or a data driven method, such as those based on dimensionality reduction techniques (for instance PCA). PCA itself being a linear transformation to the axis of greatest variation. Data driven methods have the advantage of being non-invasive, and can be performed post-acquisition. However, their main downside being that they are adversely affected by the tracer kinetics of the dynamic PET acquisition. Therefore, they are mostly limited to static PET acquisitions. This work seeks to extend on existing PCA-based data-driven motion correction methods, to allow for their applicability to dynamic PET imaging.
        \paragraph{Approach} The methods explored in this work include; a moving window approach (similar to the Kinetic Respiratory Gating method from Schleyer et al.), extrapolation of the principal component from later time points to earlier time points, and a method to score, select, and combine multiple respiratory components. The resulting respiratory traces were evaluated on $22$ data sets from a dynamic ${}^{18F}$FDG study on patients with Idiopathic Pulmonary Fibrosis. This was achieved by calculating their correlation with a surrogate signal acquired using a Real Time Position Management System.
        \paragraph{Main results} The results indicate that all methods produce better surrogate signals than when applying conventional PCA to dynamic data (for instance, a higher correlation with a gold standard respiratory trace). Extrapolating a late time point principal component produced more promising results than using a moving window. Scoring, selecting, and combining components held benefits over all other methods.
        \paragraph{Significance} This work allows for the extraction of a surrogate signal from dynamic PET data earlier in the acquisition and with a greater accuracy than previous work. This potentially allows for numerous other methods (for instance, respiratory motion correction) to be applied to this data (when they otherwise could not be previously used).
    \end{abstract}
    
    \noindent{\it Keywords\/}: PET, Dynamic PET, Surrogate Signal Extraction, PCA, Data Driven Gating
    
    \submitto{\PMB}
    
    \maketitle
    
    \input{body}

    \section*{Acknowledgments}
        This research was supported by GE Healthcare, the NIHR UCLH Biomedical Research Centre and the UCL EPSRC Centre for Doctoral Training in Intelligent, Integrated Imaging in Healthcare (i4health) grant (EP/L016478/1).
        
        The software used was partly produced by the Computational Collaborative Project on Synergistic Biomedical Imaging, CCP SyneRBI, UK EPSRC grant (EP/T026693/1).
        
        Jamie~R.~McClelland is supported by a Cancer Research UK Centres Network Accelerator Award grant (A21993) to the ART-NET consortium and a CRUK Multi-disciplinary grant (CRC 521).
    
    \printbibliography
\end{document}
